---
title: "KNN testing"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(spdep)
```





#Let's make some sample data so people can run it. First a matrix of point coordinates:
```{r}
#coords <- cbind(runif(100), runif(100)) # Example code

#coords <- cbind(ctmp[,2],ctmp[,1])
#coords <- coords[1:sub, ]
```




#Then suppose we have A measured at each of the 100 points:
# counts / weights to each coordinate:


#And some ID values which we might need:

#Then we compute the nearest neighbours (I'm doing 5 nearest for simplicity):
#Each element of knn50 is a vector of the index in coords of the neighbours:
```{r}
#A <- runif(100)
A <- ctmp[,3]
a <- A[1:sub]
#IDs = paste0("ID-",1:100)
#IDs <- 

```



So for example we can plot all the points, plot the neighbours of point 5 in green, then point 5 itself in red:
  
Then if we want to average over some measure at each of those points, here A, we can use sapply and a function that subsets from A according to those nearest elements:


giving the mean of A for the 5 nearest neighbours of each of the 100 points. Note the A value of the point itself isn't included.

## Visualize single dots:

```{r}

plot(coords)
points(coords[knn50[[5]],],col="green",pch=19)
points(coords[5,,drop=FALSE],col="red",pch=19)
sapply(1:length(knn50), function(N){mean(A[N])}) #What is this?! - Applying weights/counts to the dots? 
```



https://github.com/patrickCNMartin/VesaliusDev/blob/main/R/embeddings.R


```{r}
library("dbscan")
ctmp.knn <- kNN(coords, k = 5)
```

## Collecting data:

Barcodes and coordinates from bead_locations



```{r}
alias <- "mouse_liver_met_2_dna_200114_10"
in.path <- "~/Thesis_SpatialGenomics/SlideSeq/SlideSeq_ves_test/"
out.path <- "~/Thesis_SpatialGenomics/R_Output/"
bead <- ReadSlideSeq(paste0(in.path, alias, ".bead_locations.csv"))
#1. create a vector that contains all your barcodes.
barcodes <- rownames(bead@coordinates)
bead.coords <- bead@coordinates
coords <- cbind(bead.coords$xcoord, bead.coords$ycoord)

k <- 5
l <- length(barcodes)

ctmp.knearneigh <- knearneigh(coords, k = k)

knn50 <- knn2nb(ctmp.knearneigh, row.names = barcodes)
ctmp.nn <- ctmp.knearneigh$nn

knn50[[1]]




```

One way to think about the pooling algorithm is:
Get all the nearest neighbours for each set of coordinates (essentially Knn) associated to each barcode
1. create a vector that contains all your barcodes.



Patrick's code:
"the nearest neighbour index (nn in the first one and id in the second one) you can use that to get the barcode name (what ever each bin is called) and then group that into g number of groups."
Distance pooling array
https://github.com/patrickCNMartin/Vesalius/blob/3567500a00d5baff9274cb8e153622c5f84e15c8/R/IsolatingTerritories.R#L736-L873

.distancePooling.array <- function(img,captureRadius,minBar){

## Get knn groups of barcodes (from knn)
I don't know how to keep only unique barcodes
the length of the list should be the length / k - but I still have the full length number of groups..

There are also 14 leftovers not in any group If I understand correctly?

## Unique barcode knn list
2. Create an empty list that will grow as you assign more groups to it.
3. Take one random barcode from your barcode vector ( you can also just take the first 1 in the list) and all of its neighbours and put them in the list.
4. Remove those barcodes from your barcodes vector.
5. Repeat step 4 and 5 until all barcodes have been assigned into your list.

```{r}
knn5 <- knn50[1:l] #Just the grouped indexes (to the barcodes) - not the other attributes

bc <- barcodes
knn5.bc <- vector(mode="list", length=l) #holds unique barcodes in groups of up to k
doubles <- vector(mode="list", length=l) #Probably don't need this..
knn5.bc1 <- vector(mode="list", length=l) #unique again but without NA elements (Still null groups)

i <- 1

while(i <= l){
  j <- 1
  while(j <= k){
    ix <- knn5[[i]][j]
    b <- barcodes[ix]
    if(b %in% bc){
      knn5.bc[[i]][j] <- b
    }else{
      doubles[[i]][j] <- b
    }
    j <- j + 1
    bc <- bc[!bc %in% b]
  }
  x <- knn5.bc[[i]]
  knn5.bc1[[i]] <- x[!is.na(x)]
  i <- i + 1
}

length(bc) #leftovers

```



6. take the median x and y coordinates of all barcodes present in each group and create a new coordinate file with a new set of barcode names

```{r}
bead.x <- data.frame(bead.coords$xcoord)
bead.y <- data.frame(bead.coords$ycoord)

rownames(bead.x) <- rownames(bead@coordinates)
rownames(bead.y) <- rownames(bead@coordinates)

l <- length(knn5.bc1)
bc.g <- vector(length=l)
knn.x <- vector(length=l)
knn.y <- vector(length=l)
i <- 1

while(i <= l){
  lth <- length(knn5.bc1[[i]])
  x <- vector(length=lth)
  y <- vector(length=lth)
  j <- 1
  while(j <= lth){
    b <- knn5.bc1[[i]][j]
    x[[j]] <- bead.x[b, ]
    y[[j]] <- bead.y[b, ]
    j <- j + 1
  }
  if(length(knn5.bc1[[i]]) > 0){
    bc.g[[i]] <- knn5.bc1[[i]][1]
  }
  knn.x[[i]] <- median(x)
  knn.y[[i]] <- median(y)
  i <- i + 1
}
#knn.x %>% drop_na()
#knn.y %>% drop_na()

#knn5.bc1[[i]] <- x[!is.na(x)]

comb <- cbind(knn.x, knn.y)
as.data.frame(comb)

rownames(comb) <- bc.g

row.has.na <- apply(comb, 1, function(x){any(is.na(x))})
sum(row.has.na) #number of rows with na's to be removed


#Final output: (Should be the same length)

comb <- comb[!row.has.na,]

knn.bc <- knn5.bc1[!sapply(knn5.bc1,is.null)]


```



## Summing counts over knn-groups:

- Iterate over knn groups 
- take the counts for those barcodes for all bins from the sparsematrix
  - df - Sparse matrix vector from the DNA vesalius script - Barcodes as columns, bins as rows.
- Sum the counts for each group and each bin 

rowSums - doesn't work on groups with just 1 barcode.. 

```{r}
spMtrx <- df
i <- 1
l <- length(knn.bc)
grMtrx <- matrix(0, nrow = nrow(spMtrx), ncol = l)

while(i <= l){
  tmp <- knn.bc[[i]]
  t <- spMtrx[,tmp]
  ltmp <- length(tmp)
  if(ltmp > 1){
    grMtrx[,i] <- rowSums(t)
  }else if(ltmp == 1){
    grMtrx[,i] <- t
  }
  i <- i + 1 
}

```

# Rename columns with my new barcode names

- The first member barcode ID represents the group - stored in vector bc.g
- Same barcodes as used as rownames in vector "comb"


11. Rename the columns in your count matrix to reflect you new barcode names

```{r}

bc.comb <- rownames(comb)
colnames(grMtrx) <- bc.comb

```
