---
title: "SeqScopeVes_10um_explore"
output: html_document
---

```{r setup, include=FALSE}
library(imagerExtra)
library(imager)
library(Seurat)
library(ggplot2)
library(patchwork)
library(viridis)
library(RColorBrewer)
library(dplyr)
library(tvR)
library(sp)
library(grid)
library(parallel)
library(devtools)
source("~/R/quickload.R")
```

## Load the Seurat Object 
Ready processed - QC, normalized, DE, Annoation etc..

Save cell-types - Hep_MT or _PC etc in cells vector

```{r}
seq <-readRDS("~/group/seqScope/Liver_normal_10um_annotated.rds")

cells <- as.character(Idents(seq)) 
```


## Dimensionality Reduction
This takes time! convert complex expression Matrix to RGB colors (complex matrix -> simple table)


## Holds UMI (collapsed?), coordinates, color channel, value (of colour), slice(/tile)?
```{r}
seq <- rgbUMAP(SO = seq, conserveSparse=F)
```



## Cell plotting 
Add column with cell type (OBS 49392 observations in seq, 16464 items in cells.. 
3 times for the three different color channels)

```{r}
cells <- data.frame(seq,cells=rep(cells,times =3))
```


## Keep only Slice 2104:

```{r}
cells <- cells %>% filter(x > 60000 & x < 86000)

#Keep colour channel 1 and part of a slice:
#cells <- cells %>% filter(cc==1 & x > 6000 & x < 8500 & y<3000)
```



#Reduce number of pixels

Don't do this to the subsampled pixels for further sub-cellular analysis..

```{r}
#seq$x <- seq$x/10
#seq$y <- seq$y/10
```




## Create Colour Palette:
ter_pal: Function which picks 8 colours from the Accent Colour Panel
ter_col: Picks nuances of those colours - same number as the number of cell types
cex: Define symbol scaling size, like test and symbols, cex (10 seems good)

```{r}
ter_pal <- colorRampPalette(brewer.pal(8, "Accent")) 
ter_col <- ter_pal(length(unique(cells$cells))) 
cex <- 10
```

## UMI/Cell type plot
Plotting of all UMIs, coloured based on anotated celltype:

```{r}
an <- ggplot(cells, aes(x,y,col = cells))+ geom_point(size=1, alpha = 0.65)+
theme_void() +
scale_color_manual(values = ter_col)+
theme(legend.text = element_text(size = 12),
      legend.title = element_text(size=12),
      plot.title = element_text(size =15),
      legend.position = "left")+
guides(colour = guide_legend(override.aes = list(size=cex * 0.3)))+
labs(colour = "Cell Type", title = "Cell Types")
pdf("cellTypeLiverSeqScope.pdf", width = 6, height= 5)
print(an)
```









## Error in .checkVesalius(coordinates, sliceID, verbose) : 
image <- buildImageArray(seq, sliceID=1, resolution=30, filterThreshold=0.99, cores= 10)
tmpImage <- equalizeHistogram(image, sleft = 2.5 , sright = 2.5)
tmpImage <- regulariseImage(tmpImage, lambda = 10,invert =F)
tmpImage <- iterativeSegmentation.array(tmpImage,colDepth = 9,
  smoothIter = 10,
  method = c("box","iso"),
  acrossLevels = "mean",
  sigma = seq(1,3),
  box = 10,
  useCenter = T)
tmpImage <- isolateTerritories.array(tmpImage,minBar=5,captureRadius = 0.035)
plotTmp <- filter(tmpImage, x > 800 & x < 1700)
## what is this PAtrick??? Honestly you can do better.
## 12 hours straight of work is not an excuse for this kind of shitty code
for(i in seq_len(length(unique(plotTmp$territory)))){
  print(i)
    plotTmp$territory[plotTmp$territory == unique(plotTmp$territory)[i]] <- i
}

pdf("seqScopeTerLiver.pdf", width = 7, height=6)
g <-territoryPlot(plotTmp,cex = 15,cex.pt=1)+
   theme_void()+
   theme(legend.text = element_text(size = 12),
         legend.title = element_text(size=12),
         plot.title = element_text(size =15),
         legend.position = "right")+
  labs(title= "Vesalius - Territories", colour= "Territory nr.")
print(g)
dev.off()

pdf("seqScopeLiver_combined.pdf",width = 10,height=4.5)
print(an +g )
dev.off()

### Annotated seurat object
seq <-readRDS("~/group/seqScope/Colon_10um_annotated.rds")
cells <- as.character(Idents(seq))
#seq <- rgbPCA(SO =seq, slices=1,conserveSparse=F)
seq <- rgbUMAP(SO = seq, conserveSparse=F)


seq$x <- seq$x/10
seq$y <- seq$y/10

cells <- data.frame(seq,cells=rep(cells,times =3))
cells <- cells %>% filter(cc==1 & x > 6000 & x < 8500 & y<3000)


ter_col <- length(unique(cells$cells))
ter_pal <- colorRampPalette(brewer.pal(8, "Accent"))
ter_col <- ter_pal(ter_col)
cex <- 10

an1 <- ggplot(cells, aes(x,y,col = cells))+ geom_point(size=1, alpha = 0.65)+
theme_void() +
scale_color_manual(values = ter_col)+
theme(legend.text = element_text(size = 12),
      legend.title = element_text(size=12),
      plot.title = element_text(size =15),
      legend.position = "left")+
guides(colour = guide_legend(override.aes = list(size=cex * 0.3)))+
labs(colour = "Cell Type", title = "Cell Types")
pdf("cellTypeColonSeqScope.pdf", width = 6, height= 5)
print(an1)
dev.off()


image <- buildImageArray(seq,sliceID=1,resolution=30,filterThreshold=0.99, cores= 10)
tmpImage <- equalizeHistogram(image, sleft = 2.5 , sright = 2.5)
tmpImage <- regulariseImage(tmpImage, lambda = 10,invert =F)
tmpImage <- iterativeSegmentation.array(tmpImage,colDepth = 9,
  smoothIter = 10,
  method = c("box","iso"),
  acrossLevels = "mean",
  sigma = seq(1,3),
  box = 10,
  useCenter = T)
tmpImage <- isolateTerritories.array(tmpImage,minBar=5,captureRadius = 0.035)


plotTmp <- filter(tmpImage, x > 1600 & x < 2500 & y <1000)
for(i in seq_len(length(unique(plotTmp$territory)))){
  print(i)
    plotTmp$territory[plotTmp$territory == unique(plotTmp$territory)[i]] <- i
}

pdf("seqScopeTerColon.pdf", width = 18, height=4.5)
g1 <-territoryPlot(plotTmp,cex = 15,cex.pt=1)+
   theme_void()+
   theme(legend.text = element_text(size = 12),
         legend.title = element_text(size=12),
         plot.title = element_text(size =15),
         legend.position = "right")
print(g1)
dev.off()


pdf("seqScopeColon_combined.pdf",width = 10,height=4.5)
print(an1 +g1 )
dev.off()


### Not annotated seurat object
seq <-readRDS("~/group/seqScope/Liver_normal_Segmentation_Hepatocytes.rds")

cells <- as.character(Idents(seq))
#seq <- rgbPCA(SO =seq, slices=1,conserveSparse=F)
seq <- rgbUMAP(SO = seq, conserveSparse=F)

seq_rgbUMAP <- seq 

seq$x <- seq$x/10
seq$y <- seq$y/10

#plot(seq$x, seq$y)

cells <- data.frame(seq,cells=rep(cells,times =3))


ter_col <- length(unique(cells$cells))
ter_pal <- colorRampPalette(brewer.pal(8, "Accent"))
ter_col <- ter_pal(ter_col)
cex <- 10

#an2 <- 
ggplot(cells, aes(x,y,col = cells))+ geom_point(size=1, alpha = 0.65)+
theme_void() +
scale_color_manual(values = ter_col)+
theme(legend.text = element_text(size = 12),
      legend.title = element_text(size=12),
      plot.title = element_text(size =15),
      legend.position = "left")+
guides(colour = guide_legend(override.aes = list(size=cex * 0.3)))+
labs(colour = "Cell Type", title = "Cell Types")
pdf("cellTypeLiverCellsSeqScope.pdf", width = 6, height= 5)
print(an2)
dev.off()


image <- buildImageArray(seq,sliceID=1,resolution=30,filterThreshold=0.99, cores= 10)
tmpImage <- equalizeHistogram(image, sleft = 2.5 , sright = 2.5)
tmpImage <- regulariseImage(tmpImage, lambda = 10,invert =F)
tmpImage <- iterativeSegmentation.array(tmpImage,colDepth = 9,
  smoothIter = 10,
  method = c("box","iso"),
  acrossLevels = "mean",
  sigma = seq(1,3),
  box = 10,
  useCenter = T)
tmpImage <- isolateTerritories.array(tmpImage,minBar=5,captureRadius = 0.035)


plotTmp <- filter(tmpImage, x > 1600 & x < 2500 & y <1000)
for(i in seq_len(length(unique(plotTmp$territory)))){
  print(i)
    plotTmp$territory[plotTmp$territory == unique(plotTmp$territory)[i]] <- i
}

pdf("seqScopeTerColon.pdf", width = 18, height=4.5)
g1 <-territoryPlot(plotTmp,cex = 15,cex.pt=1)+
   theme_void()+
   theme(legend.text = element_text(size = 12),
         legend.title = element_text(size=12),
         plot.title = element_text(size =15),
         legend.position = "right")
print(g1)
dev.off()


pdf("seqScopeColon_combined.pdf",width = 10,height=4.5)
print(an1 +g1 )
dev.off()


```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
