---
title: "Segmented_SeqScope_Vesalius"
output: html_document
---

#------------------------------------------------------------------------------#
# Testing out seq scope segmented + DCT (+FFT)
#------------------------------------------------------------------------------#


```{r}
library(imagerExtra)
library(imager)
library(Seurat)
library(ggplot2)
library(patchwork)
library(viridis)
library(RColorBrewer)
library(dplyr)
library(tvR)
library(sp)
library(grid)
library(parallel)
#library(vesalius) # First version
source("~/R/quickload.R") # Dev version
```




### Load annotated seurat object
Segmented based on tissue images (or density of unspliced mRNA?) - identified nucleus position of cells
```{r}
seq.seg <-readRDS("~/group/seqScope/Liver_normal_Segmentation_Hepatocytes.rds")
cells.id <- as.character(Idents(seq.seg))
```

## Dimensionality reduction:
PCA + tSNE or UMAP
Convert to RGB colour channels

Re-Assign Cell-Types to UMIs
Repeated 3 times for the three differet colour channels.
..represented by just numbers for the segmeted dataset?

```{r}
#seq <- rgbPCA(SO =seq, slices=1,conserveSparse=F)
seq.seg <- rgbUMAP(SO = seq.seg, conserveSparse=F)
cells.seg <- data.frame(seq.seg,cells=rep(cells.id,times =3))

```
## The different tiles:
```{r}
cells.2104 <- dplyr::filter(cells.seg, grepl('2104_*', barcodes))
cells.2105 <- dplyr::filter(cells.seg, grepl('2105_*', barcodes))
cells.2106 <- dplyr::filter(cells.seg, grepl('2106_*', barcodes))
cells.2107 <- dplyr::filter(cells.seg, grepl('2107_*', barcodes))
```


## Plotting of cells and their cell types:

```{r}
cell.plot <- function(mtx, tile) {
  cex <- 10
  title <- paste("Tile:", tile, "Cell Types.")
  layout(matrix(c(1,3,2,4), 2, 2))
  an <- ggplot(mtx, aes(x,y,col = cells))+ geom_point(size=1, alpha = 0.65)+
  theme_void() +
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size=12),
        plot.title = element_text(size =15),
        legend.position = "left")+
  guides(colour = guide_legend(override.aes = list(size=cex * 0.3)))+
  labs(colour = "Cell Type", title =title)
  
  title <- paste("Tile:", tile, "Cell Types. AccentColors")
  ter_pal <- colorRampPalette(brewer.pal(8, "Accent"))
  ter_col <- ter_pal(length(unique(cells.seg$cells)))
  gg.an <- ggplot(mtx, aes(x,y,col = cells))+ geom_point(size=1, alpha = 0.65)+
  theme_void() +
  scale_color_manual(values = ter_col)+
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size=12),
        plot.title = element_text(size =15),
        legend.position = "left")+
  guides(colour = guide_legend(override.aes = list(size=cex * 0.3)))+
  labs(colour = "Cell Type", title = title)
  #pdf("cellTypeLiverSeqScope.pdf", width = 6, height= 5)
  #print(an.sect1 + an.sect2 + an1 + an2)
  print(an + gg.an)
  return(gg.an)
  #dev.off()
  }
  
cell.plot(cells.seg, 'All Overlapped')
gg.an <- cell.plot(cells.2104, '2104')
cell.plot(cells.2105, '2105')
cell.plot(cells.2106, '2106')
cell.plot(cells.2107, '2107')
```
## Select tile of interest
Decrease pixel size (optional - use for big datasets)

## Subset and Sort
Select only UMIs from tile 2104
Sort order based on colour channel and pixel location

This makes no sense - the X,Y positions and the tiles - are the barcode names really based on tiles? 
```{r}
seg <- seq.seg
seg$x <- jitter(seg$x/10,factor=0.1)
seg$y <- jitter(seg$y/10,factor=0.1)
#seq.tile <- seq.seg %>% filter(x > 6000 & x < 8600)
seg.sort <- seg[order(seg$y),]
seg.sort <- seg.sort[order(seq.sort$x),]
seg.sort <- seg.sort[order(seq.sort$cc),]
seg.2104 <- dplyr::filter(seg.sort, grepl('2104_*', barcodes))
```


##Removes any pixel overlaps 
can create problem with the x,y divided by 10 
Add jitter function before dividing to fix (get unique x,y).
```{r}
image <- buildImageArray(seg.2104,sliceID=1,resolution=30,filterThreshold=0.99, cores= 10)
im.eq <- equalizeHistogram(image, sleft = 2.5 , sright = 2.5)
im.reg <- regulariseImage(im.eq, lambda = 10,invert =F)
im.itSeg <- iterativeSegmentation.array(im.reg,colDepth = 9,
                                        smoothIter = 10,
                                        method = c("box","iso"),
                                        acrossLevels = "mean",
                                        sigma = seq(1,3),
                                        box = 10,
                                        useCenter = T)
im.terr <- isolateTerritories.array(im.itSeg,minBar=5,captureRadius = 0.035)
```




## DCT from ImagerExtra
DCT stuff:
```{r}

cimgDCT <- function(im, step) {
  ss <- subset(im, select=c(x, y, cc, value))
  layout(matrix(c(1,3,2,4), 2, 2))
  g.or <- as.cimg(ss) %>% plot(main=c(step, "Array as.cimg"))
  g <- grayscale(g.or) %>% plot(main =c(step, "Grayscale"))
  gg1 <- DCT2D(g) %>% plot(main =c(step, "DCT Transformed"))
  gg2 <- IDCT2D(gg1) %>% plot(main =c(step, "DCT Inversed"))
  print(mean((g - gg2)^2))
}

#pdf("DCT_SegmentedLiverSeqScope_comb.pdf", width = 6, height= 5)
cimgDCT(image, 'Original')
cimgDCT(im.eq, 'Histogram Equalized')
cimgDCT(im.reg, 'Regularized')
cimgDCT(im.itSeg, 'Iteratively Segmented')
```








## Min Max normalization:
normalize <- function(x) {return ((x - min(x)) / (max(x) - min(x)))}
itSeg.dct <- as.data.frame(gg1)
itSeg.dct$value <- normalize(itSeg.dct$value)
gg1 <- as.cimg(itSeg.dct) %>% plot(main="DCT Transformed, Normalized")

itSeg.dct.inv <- as.data.frame(gg2)
itSeg.dct.inv$value <- normalize(itSeg.dct.inv$value)
gg2 <- as.cimg(itSeg.dct.inv) %>% plot(main="DCT Inversed, Normalized")

## Log transformation
#logTrans <- function(mat){
#  cons <- 255/log(1+abs(max(mat)-min(mat)))
#  mat <- cons * log(mat)
#  return(mat)
#}
logTrans <- function(mat){
  mat <- mat + abs(min(mat)) + 1
  cons <- 255/log(1+abs(max(mat)-min(mat)))
  mat <- cons * log(mat)
  return(mat)
}
itSeg.dct <- as.data.frame(gg1)
itSeg.dct$value <- logTrans(itSeg.dct$value)
gg1 <- as.cimg(itSeg.dct) %>% plot(main="DCT Transformed, log-Normalized")




##FFT Imager
ss1.cim <- as.cimg(ss1)
tmp.fft <- FFT(ss1.cim)
layout(matrix(c(1,3,2,4), 2, 2))
plot(tmp.fft$real,main="Real part of the transform")
plot(tmp.fft$imag,main="Imaginary part of the transform")
sqrt(tmp.fft$real^2+tmp.fft$imag^2) %>% plot(main="Power spectrum")
# Do we get our image back?:
check <- FFT(tmp.fft$real,tmp.fft$imag,inverse=TRUE)$real #Should be the same as original
mean((check-im)^2)





## Vesalius resume:


```{r}
plotTmp <- im.terr
for(i in seq_len(length(unique(plotTmp$territory)))){
  print(i)
  plotTmp$territory[plotTmp$territory == unique(plotTmp$territory)[i]] <- i
}
#pdf("seqScopeTerLiver.pdf", width = 7, height=6)
g <- territoryPlot(plotTmp,cex = 15,cex.pt=1)+
  theme_void()+
  theme(legend.text = element_text(size = 10),
        legend.title = element_text(size=10),
        plot.title = element_text(size =12),
        legend.position = "right")+
  labs(title= "Vesalius - Territories", colour= "Territory nr.")
print(g)
#dev.off()

#pdf("seqScopeLiver_combined.pdf",width = 10,height=4.5)
print(gg.an + g)
dev.off()
```









