---
title: "SlideSeqDNA_Vesalius_Dev"
output: html_document
---


## Settings - Parameters etc

Specify dataset + Normalisation + dimensionality reduction method and other parameters

For DNA use log and PCA, or use TFIDF and LSI?

KNN duplicate vs unique - unique makes the dataset much smaller! More efficient for testing.. Duplicate makes the downstream PCA etc more robust and meaningful.

SlideDNASeq datasets:
human_colon_cancer_3_dna_191204_19
human_colon_cancer_4_dna_200114_13
human_colon_cancer_dna_4x_201027_17
mouse_liver_met_2_dna_200114_10
mouse_liver_met_1_dna_191114_05
mouse_liver_met_1_dna_191114_06


```{r}
start.time <- Sys.time()

# Specify SlideSeq DNA dataset:
alias <- "mouse_liver_met_2"
al.num <- "_dna_200114_10"



norm <- "TFIDF"             # Normalization method: "log","SCT","TFIDF","raw"
meth <- "LSI"               # Dimensionality reduction method: "PCA","PCA_L","UMAP","LSI","LSI_UMAP"
PCs <- 10                    # Number of Embeddings / dimensions / Principal Components / "LSI topics"
plotPC <- 4                 # Number of PC plots to export.

min.count <- 0
max.count <- 6

#chrsum <- "Chromosome"

k <- 25
knn.type <- "Duplicate"     # KNN method: "Unique", "Duplicate"

cores <- 5

colD <- 6                   # Column Depth (3-6)
smIter <- 10                # Smoothing Iterations
#imSegDims <- 4             # Image Segmentation Dimensions
st.meth <- "wilcox"          #Statistical method for extracting markers/bins, "wilcox", "t.test", "chisq"


param <- paste(paste0(norm, "Norm"), paste0(meth, PCs), sep = "_")

in.path <- "~/Thesis_SpatialGenomics/SlideSeq/SlideSeq_ves_test/"
out.path <- paste0("~/Thesis_SpatialGenomics/R_Output/VesaliusDev/", alias, "/")

## On MacBook:
in.path <- "~/Thesis/Thesis_SpatialGenomics/SlideSeq/SlideSeq_ves_test/"
out.path <- paste0("~/Thesis/R_Output/VesaliusDev/", alias, "/")


```


## Load packages needed

```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Thesis_SpatialGenomics/SlideSeq/SlideSeq_ves_test/")
knitr::opts_chunk$set(echo = TRUE)
library(imagerExtra)
library(Seurat)
#library(Seurat, lib.loc = "/usr/local/lib64/R-4.0.3/library") #If running on tycho RStudio Session
#library(Seurat,lib.loc="/home/hottesen/R/x86_64-pc-linux-gnu-library/4.0") #(Necessary for TFIDF, but not working on tycho RStudio)
library(Signac) #For running TFIDF
library(ggplot2)
library(patchwork)
library(viridis)
library(RColorBrewer)
library(dplyr)
library(tvR)
library(sp)
library(spdep)
library(grid)
library(gridExtra)
library(Matrix)
library(MatrixExtra)


files <- list.files("~/Thesis_SpatialGenomics/VesaliusDev_2022/Vesalius/R",pattern = ".R")
files <- paste0("~/Thesis_SpatialGenomics/VesaliusDev_2022/Vesalius/R/",files)
## If on macbook:
files <- list.files("~/Thesis/Vesalius/R",pattern = ".R")
files <- paste0("~/Thesis/Vesalius/R/",files)

for(i in seq_along(files)){
  source(files[i])
}

library(utils)
library(stats)
library(graphics)
library(grDevices)
```



## Read data from files
Vectors:
barcodes    -     holds the barcode sequences
sparse      -     Sparse count matrix, dgCMatrix

```{r}
counts <- paste0(in.path, alias, al.num, ".sparse_counts_1Mb.txt")
bead <- ReadSlideSeq(paste0(in.path, alias, al.num, ".bead_locations.csv"))

barcodes <- rownames(bead@coordinates)

sep = "\t" 
ctmp <- read.table(counts, header = TRUE, sep=sep ) %>% na.exclude()

sparse <- sparseMatrix(i= ctmp[,2],j=ctmp[,1],x=ctmp[,3]) 

# Set default point size for plotting the UMIs (KNN unique will make changes)
pt.sz <- 0.25
```




## Genomic Bins Human/Mouse:
Each bin index corresponds to 1 million bases in a specific location on a chromosome.
bin size: 1 million bases, index 1 = chromosome 1 from base 1 -> 1.000.000
Keep only the bin index numbers (1-3114. ~3.1bn bases)

Human: 3114 bins human bins. Chr2 starts at bin 251
Mouse: 2738 bins total 



```{r}

if(grepl("human",alias)){
  bin.all <- read.table(paste0(in.path,"hg19bins.txt"), header = T)
  bins <- bin.all$bin_ind
} else if(grepl("mouse",alias)){
  bin.all <- read.table(paste0(in.path,"mm10bins.txt"), header = T)
  bins <- bin.all$bin_ind
}
```



## Name column/rows in sparse matrix
Assign bin index numbers to identify the rows. 
Assign the barcodes to identify the columns

```{r}
colnames(sparse) <- barcodes
rownames(sparse) <- bins
```


Distribution:
 
max <- sort(sparse@x, decreasing = TRUE)



## Vector holders - Checkpoint

```{r}
df1 <- as.matrix(sparse) 
bead1 <- bead
param1 <- param

df <- df1
bead <- bead1
param <- param1
```

#Remove outlier counts
Expected CNA varies between 0-6? (Total deletion and double duplication?)
High outlier counts might represent mitochondrial DNA bins or repeated elements or other errors?
Mouse - outlier counts 60-506 only belong to the mitochondrial bin
Human - 


```{r}

if(exists("min.count")){
  sparse <- filterSparse(sparse, function(x) x >= min.count)
}
if(exists("max.count")){
  sparse <- filterSparse(sparse, function(x) x <= max.count)
  miax <- paste0("counts", min.count, "-", max.count)
  param <- paste(miax, param, sep="_")
}


```

## Sum counts across full chromosomes
For comparing counts over chromosomes instead of over bins.


```{r}
if(exists("chrsum")){
  sparse <- chromosomeCounts(as.matrix(sparse), bin.all)
  param <- paste("ChrSummed", param, sep="_")
}
```

## Run KNN
Sums counts over a group of neighboring barcodes
- Decide between only using each barcode once (KNN.DNA.Unique function) - Good for reducing sample size - 
  or  
  - each barcode gets a group of k nearest neighbors (KNN.DNA function) - Good for more robust downstream analyses


```{r}
bead.ls <- list()
if(exists("k")){
  if(knn.type == "Unique"){
    bead.ls <- KNN.DNA.Unique(bead, df, k)
    pt.sz <- k/15 #Adjusting point size for plotting
    param <- paste(param, paste0("knn", knn.type, k), sep="_")
  }else if(knn.type == "Duplicate"){
    bead.ls <- KNN.DNA(bead, df, k)
    param <- paste(param, paste0("knn", knn.type, k), sep="_")
  }
  sparse <- bead.ls[[1]] #Retrieve sparse matrix output from knn function
  bead <- bead.ls[[2]] #Retrieve Bead info (barcodes and coordinates) in SlideSeq assay format compatible with Seurat/Vesalius object
  rm(bead.ls)
}else{
  bead <- bead@coordinates
}

bead <- cbind(rownames(bead), bead[[1]], bead[[2]])
colnames(bead) <- c("barcodes","xcoord","ycoord")
bead <- as.data.frame(bead)
bead <- transform(bead, xcoord = as.numeric(xcoord), 
                  ycoord = as.numeric(ycoord))

sparse.df <- as.data.frame(sparse)
sparse.spmx <- Matrix(sparse, sparse = TRUE)

```

## Create Vesalius Object - Milestone
Based on the counts and bead coordinates

Beads in data.frame format - coordinates numerical
Counts either dataframe or sparse matrix.

# Dimensionality Reduction
Latent space
PCA / UMAP / tSNE (/LSI)

buildVesaliusEmbeddings - 
  - method - dimensionality reduction method, pcs - number of principal components (default 30)
  - tensorResolution - image compression ratio 0-1
  - filterThreshold - to remove outlier beads - quantile threshold at which tiles should be retained
    - 0.99 means that 99 % of barcodes and tile triangles will be retained
  
```{r}
#sparse <- sparse.df
sparse <- sparse.spmx

vesalius <- buildVesaliusObject(bead,sparse)
ves <- buildVesaliusEmbeddings(vesalius, method = meth, 
  norm = norm, pcs = PCs, tensorResolution = 0.1, cores = cores, filterThreshold =0.98)

```

## Plotting vesalius images of the PC or LSI embeddings

Imageplot with chosen number of dimensions for the different PCs

```{r}
cat( paste(Sys.time()," Plotting top", plotPC, "principal components and exporting to pdf.\n\t\t\t File located in:", out.path, "\n"))
setwd(in.path)
fname <- paste(format(Sys.time(), "%Y%m%d_%H%M%S"), "Embeddings", alias, paste0(param, ".pdf"), sep="_")
pdf(fname)
#layout(matrix(c(1,3,2,4), 2, 2))
for(i in seq(plotPC)){
  print(imagePlot(ves, dims = i))
}
dev.off()
file.rename(from = paste0(in.path, fname), to = paste0(out.path, fname))

print(paste("Runtime so far: ", difftime(Sys.time(), start.time, units="mins"), "min"))
imagePlot(ves, dims = 1)
```



## Image processing:
imageSegmentation - iterative k-means clustering
  - dims: number of dimred embeddings
  - Embedding "last": last dimred method used (if you have multiple dimred obejcts stored in your ves object)
LayerTerritory - not relevant for DNA data?

```{r}
## For testing different smoothing iterations and column depths:
#smIter <- 8 
#colD <- 3
param2 <- param
param <- param2

```


```{r}
vesImSeg <- imageSegmentation(ves,dims=1:PCs, colDepth = colD, smoothIter=smIter,smoothType="iso",sigma=2,embedding=meth)

vesTerr <- isolateTerritories(vesImSeg, trial = "last")

param <- paste(param, paste0("dims1-", PCs), paste0("colD", colD), paste0("Iter", smIter), sep="_")
print(paste("Runtime so far: ", difftime(Sys.time(), start.time, units="mins"), "min"))
```

## Export territory Object
```{r}
vesTerrDNA <- vesTerr
fname <- paste0(out.path, "Object_DNA_Territories_", alias, "_", param, ".rd")
save(vesTerrDNA, file=fname)
```


# Export territory plots to pdf
Is there a difference between the territory isolation and the territory morphing?
Also plot the first imagesegmentation

```{r}
setwd(in.path)
fname <- paste(format(Sys.time(), "%Y%m%d_%H%M%S"), "Territories", alias, paste0(param, ".pdf"), sep="_")
pdf(fname)
layout(matrix(c(1,3,2,4), 2, 2))
territoryPlot(vesTerr,cex=10,cex.pt=pt.sz)
print(imagePlot(vesImSeg))
print(getLastLog(vesTerr))
dev.off()
file.rename(from = paste0(in.path, fname), to = paste0(out.path, fname))
```



