---
title: "SlideSeqDNA"
output: html_document
---

## Preface


Set working directory for the entire markdown/notebook. - doesn't work - still need to specify path
Load packages needed


```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Thesis_SpatialGenomics/SlideSeq/SlideSeq_ves_test/")
knitr::opts_chunk$set(echo = TRUE)
library(vesalius)
library(imagerExtra)
library(Seurat)
library(ggplot2)
library(patchwork)
library(viridis)
library(RColorBrewer)
library(dplyr)
library(tvR)
library(sp)
library(spdep)
library(grid)
library(Matrix)
library(MatrixExtra)


# Standard libraries - just in case you are a command line fiend
library(utils)
library(stats)
library(graphics)
library(grDevices)
### Set seed
set.seed(1)
```

Specify dataset + Normalisation + dimensionality reduction method and parameters

```{r}
#alias <- "human_colon_cancer_4_dna_200114_13"
#alias <- "human_colon_cancer_3_dna_191204_19"
alias <- "mouse_liver_met_2_dna_200114_10"

norm <- "SeuratNorm"
#norm <- "SCTransform"
#norm <- "NoNorm"

dimred <- "3slice_PCA" #(3 PCs -> 1 slice as 3 colour channels, 3 slices = 9 top PCs)
#dimred <- "10pcs_UMAP"

in.path <- "~/Thesis_SpatialGenomics/SlideSeq/SlideSeq_ves_test/"
out.path <- "~/Thesis_SpatialGenomics/R_Output/"

#min.count <- 2
#max.count <- 6

k <- 20
#knn.type <- "Unique"
knn.type <- "Duplicate"

colD <- 3
smIter <- 20

param <- paste(norm, dimred, sep = "_")

```

In paper they do k 50 for smoothing the beads before PCA

## Read data from files
and create a sparse count matrix. (mostly 0's, bins as columns and barcodes as rows? )
To access slots of an S4 object you use @, not $


```{r}

counts <- paste0(in.path, alias, ".sparse_counts_1Mb.txt")
bead <- ReadSlideSeq(paste0(in.path, alias, ".bead_locations.csv"))




#Collect the 1000s of barcode sequences from the bead locations file
barcodes <- rownames(bead@coordinates)

#separator used in this count matrix
sep = "\t"
ctmp <- read.table(counts, header = TRUE, sep=sep ) %>% na.exclude()
# Create DG matrix: i + j - row and column index - maybe a mix of row and columns - do as.counts instead
sparse <- sparseMatrix(i= ctmp[,2],j=ctmp[,1],x=ctmp[,3]) 

#Curious about the count distribution values
max <- sort(sparse@x, decreasing = TRUE)

# Set point size for plotting the UMIs - adjusted for knn later
pt.sz <- 0.25



# invert/transpose/switch between rows and columns
#sparse <- t(sparse)

```


## Bins Human/Mouse:
Each bin index corresponds to 1million bases in a specific location on a chromosome.
bin size: 1 million bases, index 1 = chromosome 1 from base 1 -> 1.000.000
Keep only the bin index numbers (1-3114. ~3.1bn bases)

Human: 3114 bins human bins. Chr2 starts at bin 251 :
Mouse: 2738 bins total 
```{r}

if(grepl("human",alias)){
  bins <- read.table("~/Thesis_SpatialGenomics/SlideSeq/GroupSlideSeq/hg19_1Mb_bins.txt", header = T)
  bins <- bins$bin_ind
} else if(grepl("mouse",alias)){
  bins <- read.table("~/Thesis_SpatialGenomics/SlideSeq/GroupSlideSeq/mm10_1Mb_bins.txt", header = T)
  bins <- bins$bin_ind
}
```




## Name column/rows in sparse matrix
Assign bin index numbers to identify the rows. 
Assign the barcodes to identify the columns

```{r}
colnames(sparse) <- barcodes
rownames(sparse) <- bins
```

#Remove outliers - keep only counts up to 6 (CNA = 6):


```{r}

if(exists("min.count")){
  sparse <- filterSparse(sparse, function(x) x >= min.count)
}
if(exists("max.count")){
  sparse <- filterSparse(sparse, function(x) x <= max.count)
  miax <- paste0("counts", min.count, "-", max.count)
  param <- paste(miax, param, sep="_")
}


df <- as.matrix(sparse) 

#if(exists("max.count")){
#  hist(sparse@x)
#  hist(df)
#}

#sparse <- filterSparse(sparse, function(x) x >= 3)
#pct <- length(barcodes)*0.10

# Keep only barcodes with more than 100 counts total?!
#sparse.filt <- sparse[colSums(sparse > 0) >= 100, ]

```

filter out counts below 2 and above 6 
Mouse liver before: Dimensions: 2,738(bins) x 38,580(barcodes). (8,771,833 entries, 8.30% full) - remove 15.017 values

Don't know how to filter out columns (barcodes) with fewer than 100 non-zero values...?!?!?!

## Vector holders for testing..

```{r}
df1 <- df
bead1 <- bead
param1 <- param

df <- df1
bead <- bead1
param <- param1
```


## Run KNN
- Decide between only using each barcode once in KNN.DNA.Unique 
  or having many duplicate groups of barcodes (each barcode gets a group of k nearest neighbors)
  
  
length(unique(sparseDup@Dimnames[[2]])) - because I used bc.g (with duplicates) instead of bc.c (with .1 added) in the function script..
[1] 26453
length(unique(rownames(bead@coordinates)))
[1] 38580
bead rownames have been given numbers when not unique..
  

```{r}
if(exists("k")){
  if(knn.type == "Unique"){
    ls <- KNN.DNA.Unique(bead, df, k)
    sparseUq <- ls[[1]] #15737 length(unique(sparseUq@Dimnames[[2]])): 15736
    beadUq <- ls[[2]] #15736
  }else{
    ls <- KNN.DNA(bead, df, k)
    sparseDup <- ls[[1]] #38581? + 38580 length(unique(sparseDup@Dimnames[[2]])): 14612
    beadDup <- ls[[2]] #38580
  }
  param <- paste(param, paste0("knn", knn.type, k), sep="_")
  sparse <- ls[[1]] #Retrieve sparse matrix output from knn function
  bead <- ls[[2]] #Retrieve Bead info (barcodes and coordinates) in SlideSeq assay format compatible with Seurat object

  pt.sz <- k/15 #Adjusting point size for plotting
  rm(ls)
}




```


sparse <- sparseDup
bead <- beadDup

plot(beadDup@coordinates)



## Create Seurat Object - Milestone
Based on the counts

Cells - Takes the barcodes instead of cellID?
slice 1 - This is where the x+y coordinates are added from bead, identifier - barcode?
- With knn non-unique I also have non-unique cell names in the input matrix..?

Double in size for duplicate KNN? - 1,9Gb ctmp.SO

```{r}

ctmp.SO <- CreateSeuratObject(sparse, assay ="Spatial")

bead <- bead[Cells(x = ctmp.SO)]
DefaultAssay(object = bead) <- "Spatial"
ctmp.SO[["slice1"]] <- bead

```

length(unique(ctmp.SO@assays$Spatial@counts@Dimnames[[2]]))
[1] 38580


## Normalize / Find Features / Scale
Either by Seurats functions or using SCTransform or not at all
Do this make sense to do on the DNA data? 

```{r}

if(norm == "SeuratNorm"){
  ctmp <- NormalizeData(ctmp.SO)
  ctmp <- FindVariableFeatures(ctmp, nfeatures = 200)
  ctmp <- ScaleData(ctmp)
} else if(norm == "SCTransform"){
  ctmp <- SCTransform(ctmp.SO, assay = "Spatial", verbose = FALSE)
}

```

## Create image plotting function - coloured and grayscale

```{r}
## This is equal to the existing function imagePlot except This can include Title

cimg.plot <- function(image, step) {
  ss <- subset(image, select=c(x, y, cc, value))
  g <- as.cimg(ss) %>% plot(main=c(step, "Array (as.cimg)"))
}

```

## Dimensionality Reduction
Latent space
PCA / UMAP / tSNE (/Semantic Ves Dev?)

```{r}

if(grepl("UMAP",dimred)){
  pc <- sub("pcs.*", "", dimred)
  ctmp <- rgbUMAP(SO = ctmp, pcs = pc, conserveSparse=F)
} else if(grepl("PCA",dimred)){
  slice <- as.integer(substr(dimred, 1,1))
  ctmp <- rgbPCA(ctmp, slices = slice, conserveSparse = FALSE)
} else if(grepl("tSNE",dimred)){
}

```




ctmp.coords <- cbind(ctmp[[2]], ctmp[[3]])
plot(ctmp.coords)

Still have all my many bead dots..

#Image processing: - Milestone - Image array

Time consuming! (10mins - 1Gb Seurat)
Try reducing resolution to 10% - for the trial

```{r}

image <- buildImageArray(ctmp,filterThreshold=0.9975,resolution=20, cores=5)

#Keep a copy for plotting later (maybe do a plot here instead and remove the large array, print the plot later?)
image.array <- image

imagePlot(image)
cimg.plot(image, "After PCA")

```



2929734 observations of 6 variables

## Imsplit for grayscale
gray.cimg.plot <- function(image, step){
  ss.split <- subset(image.array, select=c(x, y, cc, value))
  split.im <- as.cimg(ss.split)
  split.image <- imsplit(split.im, "cc", 3)
}

```{r}


```

summary(split.im)
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
0.00000 0.01209 0.20840 0.32910 0.62528 0.99613 

> summary(split[[1]])
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
0.00000 0.01165 0.22071 0.34477 0.66823 0.99613 
> summary(split[[2]])
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
 0.0000  0.0125  0.1922  0.3140  0.6060  0.9807 
> summary(split[[3]])
   Min. 1st Qu.  Median    Mean 3rd Qu.    Max. 
0.00000 0.01209 0.20939 0.32850 0.63726 0.99252 


for(i in seq_along(split)){
  image <- equalizeHistogram(split.image[[i]], sleft = 2.5, sright=2.5,invert =T)
  image <- regulariseImage(split.image[[i]], lambda = 10, niter = 200, normalise=T)
  image <- iterativeSegmentation.array(split.image[[i]], colDepth = colD,
                                       smoothIter = smIter, method = c("iso","median"), sigma=1.5,box = 10, useCenter = T)
  image <- isolateTerritories.array(split.image[[i]], captureRadius = 0.008, minBar = 40)
  param <- paste(param, paste0("ItSegColDepth", colD), sep="_")
}

```{r}

```



```{r}
image <- equalizeHistogram(image, sleft = 2.5, sright=2.5,invert =T)
image <- regulariseImage(image, lambda = 10, niter = 200, normalise=T)
image <- iterativeSegmentation.array(image, colDepth = colD, smoothIter = smIter, method = c("iso","median"), sigma=1.5,box = 10, useCenter = T)
image <- isolateTerritories.array(image, captureRadius = 0.008, minBar = 40)
param <- paste(param, paste0("ItSegColDepth", colD), sep="_")
```

## Publish plots in pdf

```{r}
setwd(in.path)
fname <- paste(format(Sys.time(), "%Y%m%d_%H%M%S"), alias, param, "images.pdf", sep="_")
pdf(fname)
#layout(matrix(c(1,3,2,4), 2, 2))
#plot(split[[1]], main=c('Original ImageArray, 1st colour channel'))

#cimg.plot(image.array, 'Original')
#cimg.plot(image, 'Isolated Territories')


territoryPlot(image,cex=10,cex.pt=pt.sz)

dev.off()

file.rename(from = paste0(in.path, fname), to = paste0(out.path, fname))
```

