---
title: "Human DNA Colon"
output: html_document
---

## Preface


Set working directory for the entire markdown/notebook. - doesn't work - still need to specify path
Load packages needed


```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Thesis_SpatialGenomics/SlideSeq/SlideSeq_ves_test/")
knitr::opts_chunk$set(echo = TRUE)

library(imagerExtra)
library(Seurat)
library(ggplot2)
library(patchwork)
library(viridis)
library(RColorBrewer)
library(dplyr)
library(tvR)
library(sp)
library(grid)
library(Matrix)
source("~/Thesis_SpatialGenomics/RScripts_Testing/quickload.R") #Dev version of Vesalius


# Standard libraries - just in case you are a command line fiend
library(utils)
library(stats)
library(graphics)
library(grDevices)
### Set seed
set.seed(1)
```

Specify dataset + Normalisation + dimensionality reduction method and parameters

```{r}
#alias <- "human_colon_cancer_4_dna_200114_13"
#alias <- "human_colon_cancer_3_dna_191204_19"
alias <- "mouse_liver_met_2_dna_200114_10"

norm <- "SeuratNorm"
#norm <- "SCTransform"
#norm <- "NoNorm"

#dimred <- "1slice_PCA" #(3 PCs -> 1 slice as 3 colour channels, 3 slices = 9 top PCs)
dimred <- "5pcs_UMAP"
#dimred <- "1slicePCA_5pcsUMAP"

# percentage resolution in the buildimagearray - have used 20% for all before 14th March
res <- "10"


in.path <- "~/Thesis_SpatialGenomics/SlideSeq/SlideSeq_ves_test/"
out.path <- "~/Thesis_SpatialGenomics/R_Output/"
```



## Read data from files
and create a sparse count matrix. (mostly 0's, bins as columns and barcodes as rows? )


```{r}

counts <- paste0(in.path, alias, ".sparse_counts_1Mb.txt")
bead <- ReadSlideSeq(paste0(in.path, alias, ".bead_locations.csv"))



#Collect the 1000s of barcode sequences from the bead locations file
barcodes <- rownames(bead@coordinates)

#separator used in this count matrix
sep = "\t"
ctmp <- read.table(counts, header = TRUE, sep=sep ) %>% na.exclude()
# Create DG matrix: i + j - row and column index - maybe a mix of row and columns - do as.counts instead
sparse <- sparseMatrix(i= ctmp[,2],j=ctmp[,1],x=ctmp[,3]) 

#Curious about the count distribution values
max <- sort(sparse@x, decreasing = TRUE)



# invert/transpose/switch between rows and columns
#sparse <- t(sparse)

```


## Bins Human/Mouse:
Each bin index corresponds to 1million bases in a specific location on a chromosome.
bin size: 1 million bases, index 1 = chromosome 1 from base 1 -> 1.000.000
Keep only the bin index numbers (1-3114. ~3.1bn bases)

Human: 3114 bins human bins. Chr2 starts at bin 251 :
Mouse: 2738 bins total 
```{r}
if(grepl("human",alias)){
  bins <- read.table("~/Thesis_SpatialGenomics/SlideSeq/GroupSlideSeq/hg19_1Mb_bins.txt", header = T)
  bins <- bins$bin_ind
} else if(grepl("mouse",alias)){
  bins <- read.table("~/Thesis_SpatialGenomics/SlideSeq/GroupSlideSeq/mm10_1Mb_bins.txt", header = T)
  bins <- bins$bin_ind
}
```




## Name column/rows in sparse matrix
Assign bin index numbers to identify the rows. 
Assign the barcodes to identify the columns


```{r}

colnames(sparse) <- barcodes
rownames(sparse) <- bins

```

## Create Seurat Object - Milestone
Insert the Barcode sequences in the bead vector?

```{r}
ctmp.SO <- CreateSeuratObject(sparse, assay ="Spatial")

#I don't understand how this part works
bead <- bead[Cells(x = ctmp.SO)]
DefaultAssay(object = bead) <- "Spatial"
ctmp.SO[["slice1"]] <- bead
```


## Normalize / Find Features / Scale
Either by Seurats functions or using SCTransform or not at all
Do this make sense to do on the DNA data? 

```{r}

if(norm == "SeuratNorm"){
  ctmp <- NormalizeData(ctmp.SO)
} else if(norm == "SCTransform"){
  ctmp <- SCTransform(ctmp.SO, assay = "Spatial", verbose = FALSE)
}


ctmp <- FindVariableFeatures(ctmp, nfeatures = 200)
ctmp.id <- as.character(Idents(ctmp)) #Will this work?


ctmp <- ScaleData(ctmp)

```

## Create image plotting function - coloured and grayscale

```{r}
## This is equal to the existing function imagePlot except This can include Title

cimg.plot <- function(image, step) {
  ss <- subset(image, select=c(x, y, cc, value))
  g <- as.cimg(ss) %>% plot(main=c(step, "Array (as.cimg)"))
}

```

## Dimensionality Reduction
Latent space
PCA / UMAP / tSNE (/Semantic Ves Dev?)

```{r}

if(grepl("PCA",dimred)){
  slice <- as.integer(substr(dimred, 1,1))
  ctmp <- rgbPCA(ctmp, slices = slice, conserveSparse = FALSE)
} else if(grepl("UMAP",dimred) & grepl("PCA",dimred)){
  pc <- sub("pcs.*", "", dimred)
  ctmp <- rgbUMAP(SO = ctmp, pcs = pc, conserveSparse=F)
} else if(grepl("tSNE",dimred)){
}
#  slice <- as.integer(substr(dimred, 1,1))
#  ctmp <- rgbPCA(ctmp, slices = slice, conserveSparse = FALSE)
#  pc <- sub(".*PCA_", "", pc)

```

Testing to display found features..
Don't expect it to work..
Nope - they all only have 1 feature = "SeuratProject" = 1 colour

```{r}
ctmp.feat <- data.frame(ctmp,feat=rep(ctmp.id,times =3)) #will this work?!

layout(matrix(c(1,3,2,4), 2, 2))
an <- ggplot(ctmp.feat, aes(x,y,col = feat))+ geom_point(size=2, alpha = 1)+
  theme_void() +
  theme(legend.text = element_text(size = 12),
        legend.title = element_text(size=12),
        plot.title = element_text(size =15),
        legend.position = "left")+
  guides(colour = guide_legend(override.aes = list(size=10 * 0.3)))+
  labs(colour = "feat", title = "Found Features")
print(an)
```


#Image processing: - Milestone - Image array

Time consuming! (10mins - 1Gb Seurat)
Try reducing resolution to 10% - for the trial

```{r}


image <- buildImageArray(ctmp,filterThreshold=0.9975,resolution=20, cores=5)

#Keep a copy before messing up the rest
image.array <- image

```





```{r}
image <- equalizeHistogram(image.array, sleft = 2.5, sright=2.5,invert =T)
image <- regulariseImage(image, lambda = 10, niter = 200, normalise=T)

image <- iterativeSegmentation.array(image, colDepth = 6,
                                            smoothIter = 20, method = c("iso","median"), sigma=1.5,box = 10, useCenter = T)

image <- isolateTerritories.array(image, captureRadius = 0.008, minBar = 40)
```

## Publish plots in pdf
DNA - Very gray with a hint of green?

```{r}
setwd(in.path)
fname <- paste(format(Sys.time(), "%Y%m%d_%H%M%S"), alias, "VesDev", norm, dimred, res, "#Res", "images.pdf", sep="_")
pdf(fname)
layout(matrix(c(1,3,2,4), 2, 2))

cimg.plot(image.array, 'Original')


cimg.plot(image, 'Isolated Territories')

territoryPlot(image,cex=10)

dev.off()

file.rename(from = paste0(in.path, fname), to = paste0(out.path, fname))
```


```{r}

#write.csv(image, file = paste0(alias,"_ImageExport.csv"))
```

