---
title: "KNN_DNA_SlideSeq"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(spdep)
library(Seurat)
library(Matrix)
library(MatrixExtra)
```



## Start
- Make sure you have df and bead from the DNA_vesalius markdown..

- Barcodes and coordinates from bead_locations

```{r}

bead.coords <- bead@coordinates
coords <- cbind(bead.coords$xcoord, bead.coords$ycoord)

l <- length(barcodes)


```

Perform KNN:
```{r}
coords.knearneigh <- knearneigh(coords, k = k)
knn50 <- knn2nb(coords.knearneigh, row.names = barcodes)
#coords.nn <- coords.knearneigh$nn
```


## Get knn groups of barcodes (from knn)
I don't know how to keep only unique barcodes
the length of the list should be the length / k - but I still have the full length number of groups..

There are also 14 leftovers not in any group If I understand correctly?

## Unique barcode knn list
2. Create an empty list that will grow as you assign more groups to it.
3. Take one random barcode from your barcode vector ( you can also just take the first 1 in the list) and all of its neighbours and put them in the list.
4. Remove those barcodes from your barcodes vector.
5. Repeat step 4 and 5 until all barcodes have been assigned into your list.

```{r}
knn5 <- knn50[1:l] #Just the grouped indexes (to the barcodes) - not the other attributes

bc <- barcodes
knn5.bc <- vector(mode="list", length=l) #holds unique barcodes in groups of up to k
doubles <- vector(mode="list", length=l) #Probably don't need this..
knn5.bc1 <- vector(mode="list", length=l) #unique again but without NA elements (Still null groups)

i <- 1

while(i <= l){
  j <- 1
  while(j <= k){
    ix <- knn5[[i]][j]
    b <- barcodes[ix]
    if(b %in% bc){
      knn5.bc[[i]][j] <- b
    }else{
      doubles[[i]][j] <- b
    }
    j <- j + 1
    bc <- bc[!bc %in% b]
  }
  x <- knn5.bc[[i]]
  knn5.bc1[[i]] <- x[!is.na(x)]
  i <- i + 1
}

length(bc) #leftovers

```


## Find median coordinate within knn group

- iterate over groups of knn 
- stores their corresponding x and y coordinates in vector t
- Stores the median x and y for each column in new list - comb
- makes a new list of new barcode names, bc.g - the first barcode in each group

```{r}

i <- 1
l <- length(knn5.bc1)
comb <- matrix(0, nrow = l, ncol = ncol(bead.coords))
bc.g <- vector(length=l)

while(i <= l){
  tmp <- knn5.bc1[[i]]
  t <- bead.coords[tmp,]
  ltmp <- length(tmp)
  comb[i,] <- apply(t, 2, median)
  if(ltmp > 0){
    bc.g[[i]] <- tmp[[1]]
  }
  i <- i + 1 
}

rownames(comb) <- bc.g
colnames(comb) <- c("xcoord", "ycoord")

row.has.na <- apply(comb, 1, function(x){any(is.na(x))})
comb <- comb[!row.has.na,]
knn.bc <- knn5.bc1[!sapply(knn5.bc1,is.null)]

```



## Summing counts over knn-groups:

- Iterate over knn groups 
- take the counts for those barcodes for all bins from the sparsematrix
  - df - Sparse matrix vector from the DNA vesalius script - Barcodes as columns, bins as rows.
- Sum the counts for each group and each bin 

rowSums - doesn't work on groups with just 1 barcode.. 

```{r}
spMtrx <- df
i <- 1

l <- length(knn.bc)
grMtrx <- matrix(0, nrow = nrow(spMtrx), ncol = l)

while(i <= l){
  tmp <- knn.bc[[i]]
  t <- spMtrx[,tmp]
  ltmp <- length(tmp)
  if(ltmp > 1){
    grMtrx[,i] <- rowSums(t)
  }else if(ltmp == 1){
    grMtrx[,i] <- t
  }
  i <- i + 1 
}

```

# Rename columns with my new barcode names

- The first member barcode ID represents the group - stored in vector bc.g
- Same barcodes as used as rownames in vector "comb"

```{r}

bc.comb <- rownames(comb)

colnames(grMtrx) <- bc.comb
rownames(grMtrx) <- bins
knnSpMtx <- Matrix(grMtrx, sparse = TRUE)


```


## Export to files

```{r}
fname <- paste0(in.path, "knn/", alias, "_knn", k)
write.csv(comb, file = paste0(fname, ".bead_locations.csv"), row.names = TRUE, quote = FALSE)
write.table(knnSpMtx, file = paste0(fname, ".sparse_counts_1Mb.txt"), sep = "\t")



```


