---
title: "RNA from mouse Liver and Human colon"
output: html_document
---

## Preface


Set working directory for the entire markdown/notebook. - doesn't work - still need to specify path
Load packages needed


```{r setup, include=FALSE}
knitr::opts_knit$set(root.dir = "~/Thesis_SpatialGenomics/SlideSeq/SlideSeq_ves_test/")
knitr::opts_chunk$set(echo = TRUE)
library(vesalius)
library(imagerExtra)
library(Seurat)
library(ggplot2)
library(patchwork)
library(viridis)
library(RColorBrewer)
library(dplyr)
library(tvR)
library(sp)
library(grid)
library(Matrix)
library(MatrixExtra)


# Standard libraries - just in case you are a command line fiend
library(utils)
library(stats)
library(graphics)
library(grDevices)
### Set seed
set.seed(1)
```

Specify dataset + Normalisation + dimensionality reduction method and parameters

```{r}
alias <- "mouse_liver_met_2_rna_201002_04"
#alias <- "human_colon_cancer_4_rna_200102_06"

#norm <- "SeuratNorm"
norm <- "SCTransform"
#norm <- "NoNorm"

#dimred <- "6slice_PCA" #(3 PCs -> 1 slice as 3 colour channels, 3 slices = 9 top PCs)
dimred <- "30pcs_UMAP"

in.path <- "~/Thesis_SpatialGenomics/SlideSeq/SlideSeq_ves_test/"
out.path <- "~/Thesis_SpatialGenomics/R_Output/"


```



## Read data from files
and create a sparse count matrix. (mostly 0's, bins as columns and barcodes as rows? )


```{r}

counts1 <- paste0(in.path, alias, ".sparse_expression.txt")
bead1 <- ReadSlideSeq(paste0(in.path, alias, ".bead_locations.csv"))



#Collect the 1000s of barcode sequences from the bead locations file
barcodes1 <- rownames(bead@coordinates)

#separator used in this count matrix
sep = ","
ctmp1 <- read.table(counts, header = TRUE, sep=sep ) %>% na.exclude()
# Create DG matrix: i + j - row and column index - maybe a mix of row and columns - do as.counts instead
sparse1 <- sparseMatrix(i= ctmp[,2],j=ctmp[,1],x=ctmp[,3]) 

#Curious about the count distribution values
max <- sort(sparse@x, decreasing = TRUE)
hist(max)


# invert/transpose/switch between rows and columns
sparse <- t(sparse)

```






## Name columns in sparse matrix
Assign the barcodes to identify the columns
What should be row names for the RNA transcripts?!


```{r}

colnames(sparse) <- barcodes


```


## Create Seurat Object - Milestone
Insert the Barcode sequences in the bead vector?

```{r}
ctmp.SO <- CreateSeuratObject(sparse, assay ="Spatial")

#I don't understand how this part works
bead <- bead[Cells(x = ctmp.SO)]
DefaultAssay(object = bead) <- "Spatial"
ctmp.SO[["slice1"]] <- bead
```


## Normalize / Find Features / Scale
Either by Seurats functions or using SCTransform or not at all
Do this make sense to do on the DNA data? 

```{r}

if(norm == "SeuratNorm"){
  ctmp <- NormalizeData(ctmp.SO)
} else if(norm == "SCTransform"){
  ctmp <- SCTransform(ctmp.SO, assay = "Spatial", verbose = FALSE)
}


ctmp <- FindVariableFeatures(ctmp, nfeatures = 2000)


ctmp <- ScaleData(ctmp)



```

## Create image plotting function - coloured and grayscale

```{r}
## This is equal to the existing function imagePlot except This can include Title

cimg.plot <- function(image, step) {
  ss <- subset(image, select=c(x, y, cc, value))
  g <- as.cimg(ss) %>% plot(main=c(step, "Array (as.cimg)"))
}

```

## Dimensionality Reduction
Latent space
PCA / UMAP / tSNE (/Semantic Ves Dev?)

```{r}

if(grepl("UMAP",dimred)){
  pc <- sub("pcs.*", "", dimred)
  ctmp <- rgbUMAP(SO = ctmp, pcs = pc, conserveSparse=F)
} else if(grepl("PCA",dimred)){
  slice <- as.integer(substr(dimred, 1,1))
  ctmp <- rgbPCA(ctmp, slices = slice, conserveSparse = FALSE)
} else if(grepl("tSNE",dimred)){
}

```



#Image processing: - Milestone - Image array

Time consuming! (10mins - 1Gb Seurat)
Try reducing resolution to 10% - for the trial

```{r}

image <- buildImageArray(ctmp,filterThreshold=0.9975,resolution=20, cores=5)

#Keep a copy before messing up the rest
image.array <- image

```





```{r}
image <- equalizeHistogram(image.array, sleft = 2.5, sright=2.5,invert =T)
image <- regulariseImage(image, lambda = 10, niter = 200, normalise=T)

image <- iterativeSegmentation.array(image, colDepth = 6,
                                            smoothIter = 20, method = c("iso","median"), sigma=1.5,box = 10, useCenter = T)

image <- isolateTerritories.array(image, captureRadius = 0.008, minBar = 40)
```

## Publish plots in pdf

```{r}
setwd(in.path)
fname <- paste(format(Sys.time(), "%Y%m%d_%H%M%S"), alias, miax, norm, dimred, "images.pdf", sep="_")
pdf(fname)
layout(matrix(c(1,3,2,4), 2, 2))

cimg.plot(image.array, 'Original')


cimg.plot(image, 'Isolated Territories')

territoryPlot(image,cex=10)

dev.off()

file.rename(from = paste0(in.path, fname), to = paste0(out.path, fname))
```


```

# List files with the bead and count data from working directory
slideBeans - CSV files with the bead locations - barcode sequence and X vs Y coordinates
slideCounts - the expression data - here sparse expressions - counts
slideTag - the descriptive aliases that will be used

```{r}
setwd("~/SlideSeq/RNA_ves_test")
slideBeads <- list.files(pattern= "bead")
slideCounts <- list.files(pattern = "sparse")
slideTag <- gsub(".bead_locations.csv","", slideBeads)

```



```{r}
  #----------------------------------------------------------------------------#
  # Loading coordinates
  #----------------------------------------------------------------------------#
#colRNAid <- slideTag[2]
RNAid <- "mouse_liver_met_2_rna_201002_04"
#colCounts <- slideCounts[2]
counts <- "mouse_liver_met_2_rna_201002_04.sparse_expression.txt"
bead <- ReadSlideSeq("mouse_liver_met_2_rna_201002_04.bead_locations.csv")
barcodes <- rownames(bead@coordinates)
#separator used in this count matrix
sep <- ","
ctmp <- read.table(counts, header = TRUE, sep=sep ) %>% na.exclude()
# Create DG matrix: i + j - row and column index - maybe a mix of row and columns - do as.counts instead
sparse <- sparseMatrix(i= ctmp[,2],j=ctmp[,1],x=ctmp[,3])

# invert/transpose/switch between rows and columns
sparse <- t(sparse)


# human bins:
hbins <- read.table("~/SlideSeq/GroupSlideSeq/hg19_1Mb_bins.txt", header = T)
hbins <- hbins$bin_ind

# Others: (mouse)
mmbins <- read.table("~/SlideSeq/GroupSlideSeq/mm10_1Mb_bins.txt", header = T)
mmbins <- mmbins$bin_ind


colnames(sparse) <- barcodes
#rownames(sparse) <- str(mmbins)
rownames(sparse) <- ctmp[,3]

  #----------------------------------------------------------------------------#
  # Creating seurat spatial object
  # NOTE this code is taken from the Seurat source code as it does not seem that
  # Slide seq loading function are all exported
  # If this has been updated - this section can be changed accordingly
  #----------------------------------------------------------------------------#
ctmp <- CreateSeuratObject(sparse, assay ="Spatial")
bead <- bead[Cells(x = ctmp)]
  DefaultAssay(object = bead) <- "Spatial"
  ctmp[["slice1"]] <- bead
  #ctmp <- subset(ctmp, subset = nFeature_Spatial > 100 & nFeature_Spatial < 4500 & percent.mt < 5)
  #----------------------------------------------------------------------------#
  # Seuart pre-processing steps
  #----------------------------------------------------------------------------#
  ctmp <- NormalizeData(ctmp)
  ctmp <- FindVariableFeatures(ctmp, nfeatures = 2000)
  ctmp <- ScaleData(ctmp)

  #----------------------------------------------------------------------------#
  # Embed PCA loading into the RGB colour space.
  #----------------------------------------------------------------------------#
  ctmp <- rgbPCA(ctmp, slices = 3 , conserveSparse = FALSE)
  #### Remember no split ! all in the same file!!!!!
  exportRGB.csv(ctmp,file = paste0(slideTag[i],"_rgbExport.csv"),split=F)

  image <- buildImageArray(ctmp,filterThreshold=0.9975,resolution=20, cores=5)

  image <- equalizeHistogram(image,sleft = 2.5, sright=2.5,invert =T)


  image <- regulariseImage(image, lambda = 10,
                                niter = 200, normalise=T)
  image <- iterativeSegmentation.array(image, colDepth = 6,
                                            smoothIter = 20,
                                            method = c("iso","median"),
                                            sigma=1.5,box = 10,
                                            useCenter = T)
  image <- isolateTerritories.array(image,
                                          captureRadius = 0.008,
                                          minBar = 40)
  pdf(paste0(slideTag[i],"_image.pdf"))
  imagePlot(image)
  print(territoryPlot(image,cex=10))
  dev.off()
  write.csv(image, file = paste0(slideTag[i],"_ImageExport.csv"))
}
```


