---
title: "PostVesaliusTerritories"
author: "Heidi Ottesen"
date: '2022-07-06'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
in.path <- "~/Thesis_SpatialGenomics/SlideSeq/SlideSeq_ves_test/"
out.path <- paste0("~/Thesis/R_Output/VesaliusDev/RNA/", alias, "/")


files <- paste0("~/Thesis_SpatialGenomics/VesaliusDev_2022/Vesalius/R/",list.files("~/Thesis_SpatialGenomics/VesaliusDev_2022/Vesalius/R",pattern = ".R"))
## If on macbook:
files <- paste0("~/Thesis/Vesalius/R/",list.files("~/Thesis/Vesalius/R",pattern = ".R"))
for(i in seq_along(files)){
  source(files[i])
}

terr <- load(file=paste0(out.path, "Object_Territories_mouse_liver_met_2_SCTNorm_PCA30_dims1-30_colD6_Iter10_markerMethod_chisq.rd"))
```

```{r}

territoryPlot(vesTerrLay,cex=10,cex.pt=0.25)
```


```{r}
# Specify territories for comparison, seed vs query
vs <- "CloneBInnerVsOuter"
s <- c(4)           #c(8,7) or NULL
q <- c(6,13)             #c(2,3) or NULL
v <- c()
pval <- 0.05

vesMark <- extractMarkers(vesTerr,method=st.meth, seed=s, query=q, pval=pval)



vesTerrLay <- layerTerritory(vesTerr, c(s,q),morphologyFactor=c(-100, 100),layerDepth=5)

param <- paste(param2, paste0("markerMethod", st.meth), sep="_")
print(paste("Runtime so far: ", difftime(Sys.time(), start.time, units="mins"), "min"))
```





## Pick top marker bins
 - Based on lowest adjusted p value and highest, furthest from 0, Log fold change
 - With RNA we only get around 25 Genes - no need to subset..

```{r}
mark <- vesMark@DEG$DEG_Territory_Trial_1
#m.df <- as.data.frame(mark)
#padj <- slice_min(m.df, order_by=m.df$p.value.adj, n=100)
#logfc.min <- slice_min(padj, order_by=padj$logFC, n=5)
#logfc.max <- slice_max(padj, order_by=padj$logFC, n=5)
#int.genes <- bind_rows(list(logfc.min, logfc.max))
int.genes <- as.data.frame(mark)
int.genes <- int.genes[order(int.genes$p.value.adj),]
```

population[order(population$age),]


```{r}
check number of genes - acceptable to run with?
  filter by logfc as well?
```


## Plot marker bins - export to pdf
Investigate individual marker bins over the whole tissue slice.
Are some bin patterns correlating with the territory regions?

The beads are overlapping - change cex size from default 10 to less - or change alpha to change opacity/transparency




viewGeneExpression
  - why vesMorph as input? better with territory?
  - as.layer - should it show marker bin's expression/distribution per bead or overall per territory? - TRUE / FALSE
  


```{r}
layer <- "FALSE" 
for(i in seq(int.genes$genes)){
  f <- paste0(out.path, "GeneExpression/", int.genes$genes[[i]], "_", vs, "_", fname, ".png")
  png(f) 
  print(viewGeneExpression(layMark, gene=int.genes$genes[[i]], as.layer=layer, norm=F, cex=0.1, territory1=s, territory2=q))
  dev.off()
}

```

## Another bin plotting version: (One plot per pdf page)
setwd(in.path)
fname <- paste(format(Sys.time(), "%Y%m%d_%H%M%S"), "MarkerGenes", vs, alias, param, paste0("layer-", layer), ".pdf", sep="_")
pdf(fname)
for(i in seq_along(int.genes$genes)){
  print(viewGeneExpression(vesMorph, gene=int.genes$genes[[i]], as.layer=layer,norm=F,cex=0.5))
}
dev.off()
file.rename(from = paste0(in.path, fname), to = paste0(out.path, fname))

## Plot genes here in markdown:
for(i in seq_along(int.genes$genes)){
  print(viewGeneExpression(vesMorph, gene=int.genes$genes[[i]], as.layer=F,norm=F,cex=2))
}



Here I need to investigate the genes... How?!
 


```{r}

fname <- paste(format(Sys.time(), "%Y%m%d_%H%M%S"), "MarkerBins", vs, alias, param, "table.csv", sep="_")
write.csv(int.genes, file=paste0(out.path,fname))

print(paste("Runtime so far: ", difftime(Sys.time(), start.time, units="mins"), "min"))

```
```{r}

```



```{r}
print(int.genes)

viewGeneExpression(vesTerr, gene=int.genes$genes[[2]], as.layer=layer,norm=F, cex=0.1, territory1=s, territory2=q)


territoryPlot(vesTerrLay,cex=10,cex.pt=pt.sz)


```




```{r}
#Get new marker genes?


# Specify territories for comparison, seed vs query
vs <- "Sublayers"
s <- c(2)           #c(8,7) or NULL
q <- c(3)             #c(2,3) or NULL
pval <- 0.05

vesMark <- extractMarkers(vesTerr,method=st.meth, seed=s, query=q, pval=pval)


param <- paste(param2, paste0("markerMethod", st.meth), sep="_")
print(paste("Runtime so far: ", difftime(Sys.time(), start.time, units="mins"), "min"))
#Specify the new territories!
viewGeneExpression(vesTerrLay, gene=int.genes$genes[[2]], as.layer=layer,norm=F, cex=0.1, territory1=s, territory2=q)
```



## Check the log of your current vesalius object
getLastLog(vesTerr)

